Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-11 fib2.tex
more things

We consider this treatment of anyon dynamics to be a phenomenological model in that
it neglects any microscopic details of the system.
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role.
Additionally, while we model a continuum anyon theory, our analysis can be trivially reinterpreted to apply directly to lattice spin models or other systems where space is fundamentally discrete.
%\dude{but how do we relate to discrete models: toric code, string nets, etc.
%microscopic lattice models.}

@@

%We model the system on a periodic $L\times L$ square lattice $\Lambda$ of \emph{tiles} on which anyons can reside. 
%This treatment applies equally well to lattice models, where anyons 
%can only be located at a discrete set of points, 
%or continuum models, where a discretization of the manifold must 
%be chosen in order to perform charge measurements. 
%In the former case, each tile of $\Lambda$ represents a 
%lattice site, while in the latter case each tile represents 
%a measurement region. 
%For more details of an analogous model, see~\cite{Brell2013}.
\cggb{I would consider restoring the sentence that refers the reader to our previous paper for details of the phenomenological model.}

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system.
We endow this manifold with a finite set of $L\times L$ Wilson loops arranged in
a square lattice tiling $\Lambda$.
%We tile the manifold with a finite set of $L\times L$ Wilson loops arranged in
These \emph{tiles} will be the observables accesible to the error-correction procedure (\emph{decoder}).
We can use these observables to construct an idealized Hamiltonian for this system of the form
\begin{align}
	H=-\sum_{t\in \Lambda}\proj{\vac}_t\;,\label{e:hamiltonian}
\end{align}
with $\proj{\vac}_t$ the projector to charge $\vac$ at tile $t$, i.e.~the ground 
space of the model has vacuum total charge in each tile.
Given the toroidal boundary 
conditions, this space is the two-fold degenerate codespace. \cggb{Subtlety in that there could be anyons around in the ground space.}
%We identify this ground space as the codespace of our model.

Typical noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons.
It was seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study.

@@

@@

However, random noise processes are not the only dynamics that we need to consider. We 
must also consider the effect of the error-correction routine itself. This acts iteratively to fuse 
anyons on increasing length scales. While this kind of fusion would typically merge components, 
forcing us to compute dynamics of larger and larger sets of anyons,
large components are sparsely distributed
(and thus unlikely to be merged), and in addition at each length scale the total number of 
anyons present is dramatically reduced by fusion, leading to a smaller number of anyons that 
must be simulated. \cggb{is this really helping the simulation so much as it seems? is it clear that we don't still need to simulate the entire set of anyons even though they arent there anymore?}

In the regime where the combined action of noise and error-correction does not percolate
it is clear \dude{check this} that the simulation is efficient.
However, with strong enough noise the state in the fusion space will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
Despite this, we use heuristics %for finite lattice size, we use heuristics
that minimize non-local braid moves \cggb{what does minimize mean here?}, which helps to
maintain the state as a separable product beyond the
sub-percolated regime.
This enables simulation of error correction in
the regime around the error-correction threshold for linear lattice sizes up to $L=128$.

@@

%In order to keep track of the state of the system we maintain a set of disjoint curves, 
%\cggb{do they need to be directed, or is that just for the figure?} 
%representing the history of anyons on the lattice.
%These curves can also be thought of as a top-down view of the fusion tree:
%the anyons comprising the fusion tree are here displayed as points along each curve. 
%In general, the curves wind around our two 
%dimensional manifold in a haphazard way determined by the progress of the simulation.
%This creates a dynamically generated basis for the fusion space 
%that easily allows for tensor factorization of disjoint components. 
%For comparison, previous work used a single fixed basis for 
%the fusion space determined by row-major ordering of sites~\cite{Brell2013}.
\cggb{I would be an advocate for restoring the sentences that compare the dynamically generated basis here to our previous work's fixed fusion basis.}

In order to track the state of the system we maintain 
a set of disjoint directed curves, 
one for each connected component with trivial total charge.
The application of a pair-creation noise process corresponds to the 
addition of an extra curve to the lattice, supporting
two new anyons.
%along with two preferred points on it denoting the anyon locations. 
Following the noise processes, our simulation must %proceed to 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining (arbitrarily) any curves that intersect that tile, 
and then performing $R$- and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
%\dude{Heisenberg picture?}
%The observable corresponding to the tile is then
%represented in the state \dude{polish}.
%The total charge is then (probabilistically) read off the 
%relevant section of the fusion tree. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
%and the corresponding braid moves are shown in \Fref{f:syndrome}(e).
The combinatorial derivation of the algorithm and data structures used for 
this procedure will be discussed elsewhere.
%Braiding processes that encircle 
%a non-trivial loop of our (toric) manifold can also be 
%treated in an analogous way, following e.g.~\cite{Pfeifer2012}.
\cggb{what does the combinatorial comment mean?}
\cggb{why did the Pfeifer ref get commented? it should certainly go somewhere.}
%\cggb{This paragraph and the next still needs another look.}

>>>>

%\dude{but how do we relate to discrete models: toric code, string nets, etc.
%microscopic lattice models.}

@@

%We model the system on a periodic $L\times L$ square lattice $\Lambda$ of \emph{tiles} on which anyons can reside. 
%This treatment applies equally well to lattice models, where anyons 
%can only be located at a discrete set of points, 
%or continuum models, where a discretization of the manifold must 
%be chosen in order to perform charge measurements. 
%In the former case, each tile of $\Lambda$ represents a 
%lattice site, while in the latter case each tile represents 
%a measurement region. 
%For more details of an analogous model, see~\cite{Brell2013}.
%\cggb{I would consider restoring the sentence that refers the reader to our previous paper for details of the phenomenological model.}

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system.
We endow this manifold with a finite set of $L\times L$ Wilson loops arranged in
a square lattice tiling $\Lambda$.
%We tile the manifold with a finite set of $L\times L$ Wilson loops arranged in
These \emph{tiles} will be the observables accessible to the error-correction procedure (\emph{decoder}).
We can use these observables to construct an idealized Hamiltonian for this system of the form
%\begin{align}
	$H=-\sum_{t\in \Lambda}\proj{\vac}_t\;,$
%\end{align}
with $\proj{\vac}_t$ the projector to charge $\vac$ at tile $t$, i.e.~the ground 
space of the model has vacuum total charge in each tile.
Given the toroidal boundary 
conditions, this space contains the two-fold degenerate codespace.
%\cggb{Subtlety in that there could be anyons around in the ground space.}
%We identify this ground space as the codespace of our model.

Typical noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons. Such a model was analyzed in detail for Ising anyons in Ref.~\cite{Brell2013}.
It was seen that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study.

@@

We consider this treatment of anyon dynamics to be a phenomenological model in that
it neglects any microscopic details of the system.
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role.
Additionally, while we model a continuum anyon theory, our analysis can be trivially reinterpreted to apply directly to lattice spin models or other systems where space is fundamentally discrete.

@@

However, random noise processes are not the only dynamics that we need to consider. We 
must also consider the effect of the error-correction routine itself. This acts iteratively to fuse 
anyons on increasing length scales. While this kind of fusion would typically merge components, 
forcing us to compute dynamics of larger and larger sets of anyons,
large components are sparsely distributed
(and thus unlikely to be merged), and in addition at each length scale the total number of 
anyons present is dramatically reduced by fusion, leading to a smaller number of anyons that 
must be simulated.

In the regime where the combined action of noise and error-correction does not percolate
it is reasonable to expect that the simulation is efficient.
However, with strong enough noise the state in the fusion space will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
Despite this, we use heuristics %for finite lattice size, we use heuristics
that minimize braid moves, which helps to
track the state beyond the
sub-percolated regime.
This enables simulation of error correction in
the regime around the error-correction threshold for linear lattice sizes up to $L=128$.

@@

%In order to keep track of the state of the system we maintain a set of disjoint curves, 
%\cggb{do they need to be directed, or is that just for the figure?} 
%representing the history of anyons on the lattice.
%These curves can also be thought of as a top-down view of the fusion tree:
%the anyons comprising the fusion tree are here displayed as points along each curve. 
%In general, the curves wind around our two 
%dimensional manifold in a haphazard way determined by the progress of the simulation.
%This creates a dynamically generated basis for the fusion space 
%that easily allows for tensor factorization of disjoint components. 
%For comparison, previous work used a single fixed basis for 
%the fusion space determined by row-major ordering of sites~\cite{Brell2013}.
%\cggb{I would be an advocate for restoring the sentences that compare the dynamically generated basis here to our previous work's fixed fusion basis.}

In order to track the state of the system we maintain 
a set of disjoint directed curves, 
one for each connected component with trivial total charge.
Unlike previous work~\cite{Brell2013}, this creates a dynamically generated basis for the fusion space 
that easily allows for tensor factorization of disjoint components.
The application of a pair-creation noise process corresponds to the 
addition of an extra curve to the lattice, supporting
two new anyons.
%along with two preferred points on it denoting the anyon locations. 
Following the noise processes, our simulation must %proceed to 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining (arbitrarily) any curves that intersect that tile, 
and then performing $R$- and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
%\dude{Heisenberg picture?}
%The observable corresponding to the tile is then
%represented in the state \dude{polish}.
%The total charge is then (probabilistically) read off the 
%relevant section of the fusion tree. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
%and the corresponding braid moves are shown in \Fref{f:syndrome}(e).
The combinatorial derivation of the algorithm and data structures used for 
this procedure will be discussed elsewhere.
Braiding processes that encircle 
a non-trivial loop of our (toric) manifold can also be 
treated in an analogous way, following e.g.~\cite{Pfeifer2012}.
\cggb{what does the combinatorial comment mean?}
\cggb{why did the Pfeifer ref get commented? it should certainly go somewhere.}
%\cggb{This paragraph and the next still needs another look.}
